acquisition of TEs

### 2021-12-06 simulations_old_versions/Analyze_TEs.Rmd
updated paper

# load data
```{r}
cdi_ws = readRDS(here::here("Data","output_analyzeSent.Rdata"))
```

@@

## visualizing Simulation 1
simulation1 <- data_simulation1 %>%
  # convert to long data
  pivot_longer(c(WORD, DOM, NONDOM), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("DOM", "NONDOM", "WORD"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             DOM = "Panel 1A: Dominant vocabulary \n (DOM)", 
                             NONDOM = "Panel 1B: Non-dominant vocabulary \n (NONDOM)",
                             WORD = "Panel 1C: Total vocabulary \n (WORD)")) %>%
  ggplot(aes(x = number, y = TE, color = bin_BALANCE, linetype = bin_BALANCE)) +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  ylim(0, 600) +
  xlim(0, 1000) +
  labs(x = "Number of words", 
       y = "Total number of translation equivalents (TE)", 
       color = "Vocabulary balance (BALANCE)",
       linetype = "Vocabulary balance (BALANCE)",
       title = "1) Simulated data") +
  guides(color = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

@@

## Observed data: Testing Simulation 1
```{r}
Observed_plot_inputs_TE <- keepers_ws_TE %>% 
  select(baby_id, ID_testdate, age_days, number_of_te, word_vocab, TotalNW_Dom, TotalNW_NonDom, balance_vocab) %>%
  pivot_longer(c(word_vocab, TotalNW_Dom, TotalNW_NonDom), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("TotalNW_Dom", "TotalNW_NonDom", "word_vocab"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             TotalNW_Dom = "Panel 2A: Dominant vocabulary \n (DOM)", 
                             TotalNW_NonDom = "Panel 2B: Non-dominant vocabulary \n (NONDOM)",
                             word_vocab = "Panel 2C: Total vocabulary \n (WORD)")) %>%
  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) 
  ggplot(aes(x = number, y = number_of_te, color = bin_BALANCE, linetype = bin_BALANCE)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        legend.position = "bottom",
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  ylim(0, 600) +
  xlim(0, 1000) +
  labs(x = "Number of words", 
       y = "Total number of translation equivalents (TE)", 
       color = "Vocabulary balance (BALANCE)",
       linetype = "Vocabulary balance (BALANCE)",
       title = "2) Observed data") +
  guides(colour = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

@@

ggsave("Output/plot_Simulation1plusData.png", plot_Simulation1plusData,
       width = 10, height = 10)

@@

## visualizing Simulation 2
simulation2 <- data_simulation2 %>%
  pivot_longer(-c(LEARNABLE, WORD, BALANCE, bin_BALANCE), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("DOM", "NONDOM")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("NONDOM_SINGLET", "DOM_SINGLET", "TE"))) %>%
  mutate(LEARNABLE_label = case_when(LEARNABLE == 600 ~ "LEARNABLE = 600",
                                     LEARNABLE == 450 ~ "LEARNABLE = 450",
                                     LEARNABLE == 300 ~ "LEARNABLE = 300")) %>%
  group_by(LEARNABLE, LEARNABLE_label, bin_BALANCE, vocab_type) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = bin_BALANCE, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ LEARNABLE_label) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
  scale_x_discrete(labels = function(bin_BALANCE) str_wrap(bin_BALANCE, width = 10)) + 
  ylim(0, 600) +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        panel.spacing.x = unit(2, "lines")) + 
  labs(x = "\nVocabulary balance (BALANCE)", 
       y = "Number of concepts (CONCEPT)", 
       title = "Panel A: Simulated data")

@@

## Observed data: Testing Simulation 2
```{r}
plot_stacked_byLEARNABLE <- keepers_ws_TE %>%
  select(baby_id, age_months_percentile, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months \n (LEARNABLE = 244.9 - 451.9)",
                                      age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months \n (LEARNABLE = 491.8 - 604.1)",
                                      TRUE ~ "28-33 months \n (LEARNABLE = 620.4 - 638.9)")) %>%
  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "Most Balanced",
                               balance_vocab <= 0.20 ~ "Least Balanced",
                               TRUE ~ "Medium Balanced")) %>%
  mutate(BalSubset = fct_rev(BalSubset)) %>%
  arrange(desc(number_of_te)) %>%
  mutate(TE = number_of_te) %>%
  pivot_longer(-c(baby_id, age_months_percentile, balance_vocab, LEARNABLE_subset, BalSubset, word_vocab, TE), 
               names_to = "Type", values_to = "NumberOfWords") %>%
  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te"))) %>%
  mutate(Type = recode(Type, 
                       NonDom_unique_words = "Singlet in non-dominant language (NONDOM-SINGLET)", 
                       Dom_unique_words = "Singlet in dominant language (DOM-SINGLET)", 
                       number_of_te = "Translation equivalent (TE)")) %>%
  group_by(LEARNABLE_subset, BalSubset, Type) %>%
  summarise(n_words = mean(NumberOfWords)) %>%
  ggplot(aes(x = BalSubset, y = n_words, fill = Type)) +
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ LEARNABLE_subset) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  scale_x_discrete(labels = function(BalSubset) str_wrap(BalSubset, width = 10)) + 
  ylim(0, 600) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0) , 
        legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        panel.spacing.x = unit(2, "lines")) + 
  guides(fill=guide_legend(nrow=3,byrow=TRUE)) + 
  labs(x = "\nVocabulary balance (BALANCE)", 
       y = "Number of concepts (CONCEPT)",
       title = "Panel B: Observed data")

@@

## plotting the two separate models (those with WORD < 300 + those with WORD > 300)
TEprediction_percentile_facet_lessW300 <- keepers_ws_TE_simulation %>%
  mutate(W_group = if_else(word_vocab < 300, "Less than 300 total vocabulary", "More than 300 total vocabulary")) %>%
  filter(W_group == "Less than 300 total vocabulary") %>%
  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs",
       title = "Less than 300 total vocabulary") +
  xlim(0,80) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

TEprediction_percentile_facet_moreW300 <- keepers_ws_TE_simulation %>%
  mutate(W_group = if_else(word_vocab < 300, "Less than 300 total vocabulary", "More than 300 total vocabulary")) %>%
  filter(W_group == "More than 300 total vocabulary") %>%
  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs",
       title = "More than 300 total vocabulary") +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

TEprediction_percentile_facetByW300 <- 
  ggarrange(TEprediction_percentile_facet_lessW300 + rremove("x.title"), 
            TEprediction_percentile_facet_moreW300 + rremove("xy.title"), 
            ncol = 2, nrow = 1) %>%
  annotate_figure(bottom = text_grob("Number of TEs predicted by the Bilingual Vocabulary Model"))

ggsave(here("Output/plot_Simulation3_data_2WORDgroups.png"), TEprediction_percentile_facetByW300,
       height=5,width=10)
```

>>>>

# load data
```{r}
cdi_ws = readRDS(here::here("data_raw","output_analyzeSent.Rdata"))
```

@@

## visualizing Simulation 1
simulation1 <- data_simulation1 %>%
  # convert to long data
  pivot_longer(c(WORD, DOM, NONDOM), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("WORD", "DOM", "NONDOM"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             DOM = "Panel 1B: Dominant vocabulary",
                             NONDOM = "Panel 1C: Non-dominant vocabulary",
                             WORD = "Panel 1A: Total vocabulary")) %>%
  ggplot(aes(x = number, y = TE, color = bin_BALANCE, linetype = bin_BALANCE)) +
  stat_smooth(method = lm, se = F) +
#  scale_color_brewer(palette="Accent")+
  scale_color_manual(values=c("#762a83", "#abd9e9", "#a6d96a", "#fdae61", "#d53e4f")) + 
  scale_linetype_manual(values=c("solid", "solid", "solid", "solid", "solid")) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  ylim(0, 600) +
  xlim(0, 1000) +
  labs(x = "Number of words", 
       y = "Total number of translation equivalents (TE)", 
       color = "Vocabulary balance (BALANCE)",
       linetype = "Vocabulary balance (BALANCE)",
       title = "Simulated data") +
  guides(color = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

@@

## Observed data: Testing Simulation 1
```{r}
Observed_plot_inputs_TE <- keepers_ws_TE %>% 
  select(baby_id, ID_testdate, age_days, number_of_te, word_vocab, TotalNW_Dom, TotalNW_NonDom, balance_vocab) %>%
  pivot_longer(c(word_vocab, TotalNW_Dom, TotalNW_NonDom), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("word_vocab", "TotalNW_Dom", "TotalNW_NonDom"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             TotalNW_Dom = "Panel 2B: Dominant vocabulary", 
                             TotalNW_NonDom = "Panel 2C: Non-dominant vocabulary",
                             word_vocab = "Panel 2A: Total vocabulary")) %>%
  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  ggplot(aes(x = number, y = number_of_te, color = bin_BALANCE, linetype = bin_BALANCE)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#762a83", "#abd9e9", "#a6d96a", "#fdae61", "#d53e4f")) + 
  scale_linetype_manual(values=c("solid", "solid", "solid", "solid", "solid")) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        legend.position = "bottom",
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  ylim(0, 600) +
  xlim(0, 1000) +
  labs(x = "Number of words", 
       y = "Total number of translation equivalents (TE)", 
       color = "Vocabulary balance (BALANCE)",
       linetype = "Vocabulary balance (BALANCE)",
       title = "Observed data") +
  guides(colour = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

@@

ggsave("Output/BU_plot_Simulation1plusData.png", plot_Simulation1plusData,
       width = 10, height = 10)

@@

## visualizing Simulation 2
simulation2 <- data_simulation2 %>%
  pivot_longer(-c(LEARNABLE, WORD, BALANCE, bin_BALANCE), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("DOM", "NONDOM")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("NONDOM_SINGLET", "DOM_SINGLET", "TE"))) %>%
  mutate(LEARNABLE_label = case_when(LEARNABLE == 600 ~ "LEARNABLE = 600",
                                     LEARNABLE == 450 ~ "LEARNABLE = 450",
                                     LEARNABLE == 300 ~ "LEARNABLE = 300")) %>%
  group_by(LEARNABLE, LEARNABLE_label, bin_BALANCE, vocab_type) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = bin_BALANCE, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ LEARNABLE_label) +
  scale_fill_manual(values = c("#bdbdbd", "#737373", "#252525")) +
  scale_x_discrete(labels = function(bin_BALANCE) str_wrap(bin_BALANCE, width = 10)) + 
  ylim(0, 600) +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        panel.spacing.x = unit(2, "lines")) + 
  guides(fill=guide_legend(nrow=3,byrow=TRUE)) + 
  labs(x = "\nVocabulary balance (BALANCE)", 
       y = "Number of concepts (CONCEPT)", 
       title = "Panel A: Simulated data")

@@

## Observed data: Testing Simulation 2
```{r}
plot_stacked_byLEARNABLE <- keepers_ws_TE %>%
  select(baby_id, age_months_percentile, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months \n (LEARNABLE = 244.9 - 451.9)",
                                      age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months \n (LEARNABLE = 491.8 - 604.1)",
                                      TRUE ~ "28-33 months \n (LEARNABLE = 620.4 - 638.9)")) %>%
  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "Most Balanced",
                               balance_vocab <= 0.20 ~ "Least Balanced",
                               TRUE ~ "Medium Balanced")) %>%
  mutate(BalSubset = fct_rev(BalSubset)) %>%
  arrange(desc(number_of_te)) %>%
  mutate(TE = number_of_te) %>%
  pivot_longer(-c(baby_id, age_months_percentile, balance_vocab, LEARNABLE_subset, BalSubset, word_vocab, TE), 
               names_to = "Type", values_to = "NumberOfWords") %>%
  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te"))) %>%
  mutate(Type = recode(Type, 
                       NonDom_unique_words = "Singlet in non-dominant language (NONDOM-SINGLET)", 
                       Dom_unique_words = "Singlet in dominant language (DOM-SINGLET)", 
                       number_of_te = "Translation equivalent (TE)")) %>%
  group_by(LEARNABLE_subset, BalSubset, Type) %>%
  summarise(n_words = mean(NumberOfWords)) %>%
  ggplot(aes(x = BalSubset, y = n_words, fill = Type)) +
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ LEARNABLE_subset) +
  scale_fill_manual(values = c("#bdbdbd", "#737373", "#252525")) +
  scale_x_discrete(labels = function(BalSubset) str_wrap(BalSubset, width = 10)) + 
  ylim(0, 600) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0) , 
        legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        panel.spacing.x = unit(2, "lines")) + 
  guides(fill=guide_legend(nrow=3,byrow=TRUE)) + 
  labs(x = "\nVocabulary balance (BALANCE)", 
       y = "Number of concepts (CONCEPT)",
       title = "Panel B: Observed data")

@@

## plotting the two separate models (those with WORD < 300 + those with WORD > 300)
TEprediction_percentile_facet_lessW300 <- keepers_ws_TE_simulation %>%
  mutate(W_group = if_else(word_vocab < 300, "Less than 300 total vocabulary", "More than 300 total vocabulary")) %>%
  filter(W_group == "Less than 300 total vocabulary") %>%
  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs",
       title = "Less than 300 total vocabulary") +
  xlim(0,80) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  theme_minimal(base_size = 18) +
  theme(plot.title = element_text(hjust = 0.5))

TEprediction_percentile_facet_moreW300 <- keepers_ws_TE_simulation %>%
  mutate(W_group = if_else(word_vocab < 300, "Less than 300 total vocabulary", "More than 300 total vocabulary")) %>%
  filter(W_group == "More than 300 total vocabulary") %>%
  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs",
       title = "More than 300 total vocabulary") +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  theme_minimal(base_size = 18) +
  theme(plot.title = element_text(hjust = 0.5))

TEprediction_percentile_facetByW300 <- 
  ggarrange(TEprediction_percentile_facet_lessW300 + rremove("x.title"), 
            TEprediction_percentile_facet_moreW300 + rremove("xy.title"), 
            ncol = 2, nrow = 1) %>%
  annotate_figure(bottom = text_grob("Number of TEs predicted by the Bilingual Vocabulary Model", size= 18))

ggsave(here("Output/BU_plot_Simulation3_data_2WORDgroups.png"), TEprediction_percentile_facetByW300,
       height=5,width=10)
```
